global-model (ODE)

### 2015-01-30 documentation/interactive_nullmodel.Rmd
labelling nullmodel document
ODE of global model, without spatially-explicit terms (Shiny)

---
title: "null-model ode"
author: "Florian Schneider"
date: "Thursday, November 20, 2014"
output: html_document
runtime: shiny
---

## original model (Kéfi et al 2007)

Running the pair approximation model of Kéfi et al 2007 TPB dynamically.

@@

### plant and soil parameters
```{r, echo=FALSE}

inputPanel(
  sliderInput("m", label = "intrinsic mortality:",
              min = 0.0, max = 1, value = 0.05, step = 0.01),
  sliderInput("r", label = "max. growth rate:",
              min = 0.0, max = 3, value = 1, step = 0.01),
  sliderInput("b", label = "environmental quality:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("K", label = "carrying capacity:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("a", label = "search efficiency:",
              min = 0.0, max = 1, value = 0.6, step = 0.01), 
  sliderInput("h", label = "handling time:",
              min = 0.0, max = 1000, value = 100, step = 1), 
  sliderInput("L", label = "livestock units:",
              min = 0.0, max = 100, value = 10, step = 1)
)

@@

 parms <- list()
  parms$m <- as.numeric(input$m)
  parms$r <- as.numeric(input$r)
  parms$b <- as.numeric(input$b) 
  parms$K <- as.numeric(input$K) 
  parms$a <- as.numeric(input$a) 
  parms$h <- as.numeric(input$h) 
  parms$L <- as.numeric(input$L)

  plot(rho,  with(parms,  m * rho + (a * rho * L)/(1 + a * h * rho)), ylab = "growth", type = "l", ylim = c(0,0.5))
  lines(rho, with(parms, r* b* rho * (1 - rho/K) ), col = "green3" )

@@

d_rho <- function(rho, parms, z = 4) { 
  with(parms, r* b* rho * (1 - rho/K) - m * rho - (a * rho * L)/(1 + a * h * rho) )  
}

@@

 parms <- list()
  parms$m <- as.numeric(input$m)
  parms$r <- as.numeric(input$r)
  parms$b <- as.numeric(input$b) 
  parms$K <- as.numeric(input$K) 
  parms$a <- as.numeric(input$a) 
  parms$h <- as.numeric(input$h) 
  parms$L <- as.numeric(input$L)

# running the ode-solver
runmodel <- ode(y = input$rho_1, func = odesys, times = 1.1^(1:53), parms = parms)

>>>>

---
title: "global-model (ODE)"
author: "Florian Schneider"
date: "Thursday, January 30, 2015"
output: html_document
runtime: shiny
---

## developping the null model

Non-spatial global model solved as ordinary differential equations. It includes global accelerating and decelerating feedbacks on mortality and grazing.

@@

### growth parameters

The basic assumption is logistic plant growth, i.e. growth goes to zero as the carrying capacity is approached (decelerating effect).

As an accelerating feedback we add water runoff, which reduces growth at low cover.

inputPanel(
  sliderInput("r", label = "max. growth rate:",
              min = 0.0, max = 3, value = 1, step = 0.01),
  sliderInput("b", label = "environmental quality:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("K", label = "carrying capacity:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("alpha", label = "water runoff:",
              min = 0.0, max = 1, value = 0, step = 0.01)
)

### mortality parameters

The basic assumption is a constant intrinsic mortality of plants and a type II functional response, i.e. grazing mortality saturates with cover (decelerating effect).

As an additional accelerating effect we assume the search efficiency to be lowest at low cover and increase with cover, turning the function into a type III functional response.

inputPanel(
  sliderInput("m", label = "intrinsic mortality:",
              min = 0.0, max = 1, value = 0.05, step = 0.01),
  sliderInput("a", label = "search efficiency:",
              min = 0.0, max = 1, value = 0.6, step = 0.01), 
  sliderInput("h", label = "handling time:",
              min = 0.0, max = 1000, value = 100, step = 1), 
  sliderInput("L", label = "livestock units:",
              min = 0.0, max = 100, value = 10, step = 1),
  sliderInput("q", label = "hill coefficient:",
              min = 0.0, max = 3, value = 1, step = 0.01)
)

@@

 parms <- list()
  parms$m <- as.numeric(input$m)
  parms$r <- as.numeric(input$r)
  parms$b <- as.numeric(input$b) 
  parms$K <- as.numeric(input$K) 
  parms$a <- as.numeric(input$a) 
  parms$h <- as.numeric(input$h) 
  parms$L <- as.numeric(input$L) 
  parms$alpha <- as.numeric(input$alpha) 
  parms$q <- as.numeric(input$q)

  plot(rho,  with(parms,  m * rho + (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q))), ylab = "growth", type = "l", ylim = c(0,0.5))
  lines(rho, with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) ), col = "green3" )

@@

d_rho <- function(rho, parms, z = 4) { 
  with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) - m * rho - (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q)) )  
}

@@

 parms <- list()
  parms$m <- as.numeric(input$m)
  parms$r <- as.numeric(input$r)
  parms$b <- as.numeric(input$b) 
  parms$K <- as.numeric(input$K) 
  parms$a <- as.numeric(input$a) 
  parms$h <- as.numeric(input$h) 
  parms$L <- as.numeric(input$L) 
  parms$alpha <- as.numeric(input$alpha) 
  parms$q <- as.numeric(input$q)

# running the ode-solver
runmodel <- ode(y = input$rho_1, func = odesys, times = 1.1^(1:50), parms = parms)
